---
title: "Piece_6"
author: "Benjamin Egan"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse) 
library(ggplot2)
library(dsbox)
library(mosaicData) 


#datasets
plastic_waste <- read.csv("data/plastic-waste.csv")
nobel <- read_csv("data/nobel.csv")
states <- read_csv("data/states.csv")
staff <- read_csv("data/instructional-staff.csv")
```

# Useful Lines of Code

Up until this point in time, I have been looking back through each of my labs to pull and adapt lines of code. In this piece, I will be walking through all of my labs and extracting the helpful functions I have been repeatedly using.

## Filter()

```{r filter}

true_plastic <- plastic_waste %>%
  filter(plastic_waste_per_cap < 3, mismanaged_plastic_waste < 3000000, total_pop < 250000000, coastal_pop < 150000000)


```

Filter() is aptly names, as it filters out certain data points. In this lab (https://datascience4psych.github.io/DataScience4Psych/lab02.html), I used filter to exclude certain countries from a visual.


## Mutate()

Mutate() has been used across the class in different capacities. I've included examples below on how I have used mutate used:

```{r mutate lab 3}

nobel_living <- nobel %>%
  mutate(country_us = if_else(country == "USA", "USA", "Other")
  )
```
Here mutate was used to create a binary categorical variable. It kept USA as USA and changed everything else to "other"

#### List - c(a,b,c,etc.)
```{r mutate lab 4}

lq_USA <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE","PU","SL") ~ "Mexico",
    state == "FM" ~ "Honduras"
  ))

```
This is a more complex version of mutate(). Here I created a new country variable by taking all the cities and recoding them into their countries. The list function [c()] is useful for when you have multiple items you are attempting to recode.

## Wide to Long transformation - pivot_longer()

```{r wide to long}

staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year")

```
pivot_longer() is useful for changing datasets from wide to long. The first argument is the data. The second argument, cols, is where you specify which columns to pivot into longer format – in this case all columns except for the faculty_type. The third argument, names_to, is a string specifying the name of the column to create from the data stored in the column names of data – in this case year.

## Input - tibble()
string input should go in qutations while numbers would be input the same.
```{r tibble}

Name_of_table <- tribble(
  ~IV1, ~IV2, ~IV3,
  "score","score","score"
)

#example
Kansas_Mask <- tribble(
  ~date, ~count, ~mask_status,
  "7/12", 25.5, "mask",
  "7/13", 19.8, "mask",
  "7/14", 19.7, "mask",
  "7/15", 20.5, "mask",
  "7/12", 9.8, "no mask",
  "7/13", 9.2, "no mask",
  "7/14", 9.5, "no mask",
  "7/15", 9.8, "no mask"
)
```


## Functions - function()

```{r function}

function_name <- function(thing_you_change){
  
  #function pulled from lab 8 scripts
   page <- read_html(url)
  
  # scrape titles
  titles <- page %>%
    html_nodes(".iteminfo") %>%
    html_node("h3 a") %>%
    html_text() %>%
    str_squish()
  
  # scrape links
  links <- page %>%
    html_nodes(".iteminfo") %>%
    html_node("h3 a") %>%
    html_attr("href") %>%
    str_replace(".", "collections.ed.ac.uk")
  
  
  # scrape artists 
  artists <- page %>%
    html_nodes(".iteminfo") %>%
    html_node(".artist") %>%
    html_text() %>%
    str_squish()
  
  # create and return tibble
  tibble(
    title = titles,
    artist = artists,
    link = links
  )
  
}

```

## ggplot and graphing

A goal of this course was to get students comfortable graphing data. was to get students familiar with different ways to plot data using ggplot(). Below I include  different types of plots you can create using ggplot():

#### Generally useful functions

theme_bw() removes all gray from the graph.
labs() adds titles and labesl to the graphs.
scale_color_manual() allows for you to manually set colors.
fct_reorder(data, variable) allows for you to reverse the order of a variable.

### Histogram - geom_histogram()
binwidth changes how many observations fit within each bar.
```{r histogram}
true_plastic %>%
ggplot(aes(
  x = plastic_waste_per_cap)) +
  geom_histogram(color = "white", binwidth = 0.2) +
  theme_bw()+
  labs(
      x = "Plastic waste per captia by country",
      y = "Amount",
      title = "Plastic Waste"
  )
```

### Density plot - geom_density()
Alpha changes the opacity of the graph.
```{r density plot}
true_plastic %>%
ggplot(aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )) +
  geom_density(alpha = 0.7) +
  theme_bw()+
  labs(
      x = "Plastic waste per captia",
      y = "amount",
      title = "Density map of Plastic Waste by country"
  )
```


### Boxplot - geom_boxplot()
```{r boxplot}
true_plastic %>%
ggplot(aes(
    x = continent,
    y = plastic_waste_per_cap
  )) +
  geom_boxplot() +
  theme_bw()+
labs(
      x = "Plastic waste per captia",
      y = "amount",
      title = "Side by side of Plastic Waste by Continent"
  )
```

### Violin plot - geom_violin()
```{r violin plot}
true_plastic %>%
ggplot(aes(
    x = continent,
    y = plastic_waste_per_cap
  )) +
 geom_violin() +
  theme_bw()+
  labs(
      x = "Continent",
      y = "Plastic Waste",
      title = "Violin map of Plastic Waste by Continent"
  )
```

### Scatterplot - geom_point
```{r scatterplot}
true_plastic %>%
ggplot(aes(
    x = mismanaged_plastic_waste,
    y = plastic_waste_per_cap,
    color = continent
  )) +
  geom_point() +
  theme_bw()+
labs(
      x = "Mismanaged Plastic Waste",
      y = "Plastic Waste",
      title = "Plastic Waste x mismanaged plastic waste by Continent",
      color = "Continent"
  )
```

### Break down graphs by levels of a variable - facet_wrap()

Place the variable you want to be represented by the different graphs inside: (~variable of interest). Something to note is that the axes You can input (scale = "free") to un
```{r set up for facet, include = FALSE}
nobel_living <- nobel %>%
 filter(country != "NA" & gender != "org" & is.na(died_date))

nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r facet}
nobel_living_science %>%
ggplot(aes(
    x = country_us,
  )) +
  facet_wrap(~category)+
  theme_bw()+
  geom_bar()+
  coord_flip()+
  labs(
      x = "Country",
      y = "Number of Prize Winners",
      title = "Relationship Between Prize Winner and Country"
  )

```

### Making a line graph - geom_line

```{r graph, fig.width = 8}
staff_long %>%
  ggplot(aes(
    x = year,
    y = value,
    group = faculty_type,
    color = faculty_type
  )) +
  geom_line()+
  theme_bw()+
  labs(
    x = "Year",
    y = "Percent of Total Instructional Staff",
    title = "Trends in Instructional Staff Employee Status",
    color = "Faculty Status"
  )
```

### Changing parts of your visual - theme()
allows you to change axes, legends, the visuals, etc. You can use angle to change the angle of the axes
```{r theme}
data("Whickham")
Whickham %>%
ggplot(aes(
   x=age,
   fill = fct_rev(smoker)
   )) + 
    geom_histogram(color = "black")+
  theme_bw()+
  labs(
    x = "Age",
    y = "",
    title = "Smokers and non smokers at baseline",
    fill = "Are they a smoker?"
  )+
    theme(plot.title = element_text(size = 20),
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 15),
         axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.5)
        )
```

### Add text to a graph - geom_col()


### Handling categorical data - geom_jitter()

``` {r geom_jitter}


attractive_score <- evals %>%
  ggplot(aes(
    x = score,
    y = bty_avg
  ))+
  labs(
    x = "Score",
    y = "Beauty rating",
    title = "Relationship between professor rating and attractiveness",
    subtitle = "Average professor evaluation score: (1) very unsatisfactory - (5) excellent"
  )

attractive_score + geom_smooth(se = FALSE, color = "black", span = 1)

attractive_score + geom_point() + geom_smooth(se = FALSE, color = "black", span = .5)

attractive_score + geom_jitter(width = .05) + geom_smooth(se = FALSE, color = "black", span = .5)


```


### Line of best fit - geom_smooth()


## Other useful functions

![Caption](File Name) - Lets you embed an image